Peer Assessment 1 for Reproducible research 
========================================================

## Submitted on May 16, 2014 by M. S. Ravi



Reading and cleaning up data:

```{r}
activity<-read.csv("activity.csv")
head(activity)
```

Creating a new table which only has the rows without NA in the steps column.

```{r good data}
good<-complete.cases(activity)
goodactivity<-activity[good,]
```
Initializing a new array to store data about the steps per day and initializing the day.
```{r initialize stepsinday}
dates<-numeric(length=100)
steps<-numeric(length=100)
stepsinday <- data.frame(dates,steps)
day<-as.Date("2012-10-01")
```
Updating the table stepsinday, so that it counts the number of steps in each day.
```{r updating stepsinday}
j<-1
for (i in 1:nrow(goodactivity)){
      #if we are in the same day, we won't change the day, 
      # but just add the number of steps
      if (as.Date(goodactivity[i,2])==day){
            stepsinday[j,2]<-stepsinday[j,2]+goodactivity[i,1]
            }
      #if the day has changed, we go to the next row in stepsinday and start
      #ading the number of steps
      else {j<-j+1
            day<-as.Date(goodactivity[i,2])
            stepsinday[j,1]<-day
            stepsinday[j,2]<-stepsinday[j,2]+goodactivity[i,1]
            }
      }
```
Since we intialized a larger table than needed, we have a buch of irrelevant zeros in the table stepsinday. We remove them by first converting the zeros to NA, and then getting rid of the NAs.

```{r}
stepsinday[stepsinday==0] <- NA
stepsinday<-na.omit(stepsinday)
```
Note that the days are in the numeric format, of the number of days from Jan.1, 1970. Couldn't do too much about this, so I left it alone. I will now plot the histogram of the steps per day and get the mean and the median of this column.
```{r}
hist(stepsinday[,2],main="Histogram of steps per day", xlab="Number of Steps in a day",ylab="frequency")
 mean(stepsinday[,2])
median(stepsinday[,2])
```

